---
title: "Thoughts on Storks - Model selection and averaging"
author: "Lizzie Jones"
date: "26/07/2021"
output: html_document
---


```{r packages, messages = FALSE, include=FALSE}
# Assuming all packages are installed. If not use 'install.packages' e.g. install.packages("likert")


# Key packages for data wrangling
wrangling <- c("dplyr","tidyverse","purr","magrittr", "Hmisc",
             "data.table","plyr","tidyr","tibble","reshape2")
lapply(wrangling, require, character.only = TRUE) 

# Useful survey analysis packages
survey <- c("likert","careless", "survey")
lapply(survey, require, character.only = TRUE) 

# Useful packages for statistics
stats <- c("stats","ggpubr","lme4","MASS","car","psych","corrplot",
                   "MuMIn","glmmTMB","nlme","DHARMa", "glmmTMB")
lapply(stats, require, character.only = TRUE) 

# Useful packages for text analysis
text <- c("tm","koRpus","textstem", "tidytext","text2vec","lexicon","SentimentAnalysis","SnowballC",
           "wordcloud", "sentimentr", "udpipe", "wordnet") # "qdap")
lapply(text, require, character.only = TRUE)

# Favourite data visualisation packages
vis <- c("ggvis","htmlwidgets","maps", "lattice","ggmap","ggplot2","plotly",
            "RColorBrewer", "sjPlot", "ggrepel", "stargazer")
lapply(vis, require, character.only = TRUE) 


```

```{r setup, messages = FALSE, include=FALSE}

# Load in working  directory and datasets
setwd("~/Documents/Thoughts_on_Storks/Working_directory")
# Load in the data
original_data <- read.csv("Stork_MainDataset.csv", header = TRUE, stringsAsFactors=TRUE)
all_data <- read.csv("Stork_Dataset_Radapted.csv", header = TRUE, stringsAsFactors=TRUE)

# Load in cleaned dataset
final_data <- read.csv("WSP_R_FINAL_dataset2.csv", header = TRUE, stringsAsFactors=TRUE)
nrow(final_data)

```

```{r code palette, include = FALSE, messages = FALSE}
## Creating a custom colour palette
theme_set(theme_minimal())
wsp_colors <- c(
  `turquoise`  = "#5ab4ac",
  `light_tur`  = "#acd9d5",
  `grey`       = "#e5e5e5",
  `light_gold` = "#ebd9b2",
  `gold`       = "#d8b365")

wsp_cols <- function(...) {
  cols <- c(...)

  if (is.null(cols))
    return (wsp_colors)

  wsp_colors[cols]
}

# Create two custom palettes using the wsp_cols()called 'likert' and 'light'
wsp_palettes <- list(
  `likert`  = wsp_cols("turquoise", "light_tur", "grey", "light_gold", "gold"),
  `light`  = wsp_cols("light_tur", "grey", "light_gold")
)

# A function to access and interpolate 
wsp_pal <- function(palette = "likert", reverse = FALSE, ...) {
  pal <- wsp_palettes[[palette]]
  if (reverse) pal <- rev(pal)
  colorRampPalette(pal, ...)
}
# Check the 5-colour palette
wsp_pal("light")(10)


### Scales for ggplot2 plots - either fill (for geom_bar etc.) or colour (for geom_point etc.)
## COLOUR
scale_color_wsp <- function(palette = "likert", discrete = TRUE, reverse = FALSE, ...) {
  pal <- wsp_pal(palette = palette, reverse = reverse)

  if (discrete) {
    discrete_scale("colour", paste0("wsp_", palette), palette = pal, ...)
  } else {
    scale_color_gradientn(colours = pal(256), ...)
  }
}

## FILL
scale_fill_wsp <- function(palette = "likert", discrete = TRUE, reverse = FALSE, ...) {
  pal <- wsp_pal(palette = palette, reverse = reverse)

  if (discrete) {
    discrete_scale("fill", paste0("wsp_", palette), palette = pal, ...)
  } else {
    scale_fill_gradientn(colours = pal(256), ...)
  }
}

```


## Quantitative analysis/modelling 

**Quantitative data analysis:** 
Descriptive and statistical - to understand variation in respondent’s awareness, knowledge and attitudes towards white storks and their reintroduction.  

**Methods plan**

* GLM approach + model selection and averaging
    + Anderson, D. and Burnham, K., 2004. Model selection and multi-model inference. Second. NY: Springer-Verlag, 63(2020), p.10.
    + Burnham, K.P., Anderson, D.R. and Huyvaert, K.P., 2011. AIC model selection and multimodel inference in behavioral ecology: some background, observations, and comparisons. Behavioral ecology and sociobiology, 65(1), pp.23-35.    
* Compare OverallAttitudeScore to "Q15. Do you support WS reintro (yes/no)"


### Possible predictor variables

#### Factor variables

* Age (collapse further?) 
* Gender (female / male) 
* Urban / suburban / rural 
* Highest education (collapse – e.g. degree; below degree) 
* Occupation (use? If so, would need to collapse! unemployed; retired; potentially pool responses except for those who answered “environment, nature & wildlife”) 
* Visited Knepp (yes / no) 
* Time spent in nature 
* Member of conservation/environmental organisation (quite a few people listed RSPB) 
* Awareness 
* Heard of white stork before taking this survey?  
* Heard of white stork project / reintroduction effort? 

#### Numeric variables

* Contact and connection with nature; general environmental attitudes and behaviour 
* Nature Connection Index (composite score) 
* Environmental concern (composite score) 
* General attitude towards birds (composite score) 



```{r create model data, messages = FALSE, warning=FALSE}

# First, rename that REALLY long column (Q22)
names(final_data)[names(final_data) == "Q22....Are.you.a.member.of.any.environmental..wildlife.or.conservation.organisations."] <- "Q22_env_org_member"

final_data <- final_data %>% mutate(Q8_Seen =
                     case_when(Q8_wild_seen == 1L ~ "Wild", 
                               Q8_wild_seen == 1L & Q8_captivity_seen == 1L ~ "Wild", 
                               Q8_wild_seen == 0L & Q8_captivity_seen == 1L ~ "Captivity",
                               Q8_wild_seen == 0L & Q8_captivity_seen == 1L & Q8_pictures_video == 1L ~ "Captivity",
                               Q8_wild_seen == 1L & Q8_captivity_seen == 1L & Q8_No == 1L ~ "No/Not sure",
                               Q8_wild_seen == 0L & Q8_captivity_seen == 1L & Q8_No == 1L ~ "No/Not sure",
                               Q8_wild_seen == 1L & Q8_captivity_seen == 1L & Q8_NotSure == 1L ~ "No/Not sure",
                               Q8_wild_seen == 0L & Q8_captivity_seen == 1L & Q8_NotSure == 1L ~ "No/Not sure",
                               Q8_wild_seen == 0L & Q8_captivity_seen == 1L & Q8_NotSure == 1L & Q8_NotSure == 1L ~ "No/Not sure",
                               Q8_wild_seen == 0L & Q8_captivity_seen == 0L & Q8_pictures_video == 1L ~ "No/Not sure",
                               Q8_wild_seen == 0L & Q8_captivity_seen == 0L & Q8_pictures_video == 0L ~ "No/Not sure"))

# Select all possible predictor vars
model_data <- final_data %>% 
  dplyr::select(UniqueID_all, OverallAttitudeScore, SiteLocal, SurveyType,
         Age_short, Gender, Area_type, Education_short, Occupation_short_clean,
         Q14.5_agreement_score, Q8_Seen,
         Q27_Knepp_visit, Q18_exp_nature, Q15_WSP_support,
         Q1_aware_stork, Q9_heard, KnowledgeScore, Q22_env_org_member, 
         NCI, ProCoBS, BirdInterestScore, EnvConcern.score)
nrow(model_data)
model_clean <- model_data[!is.na(model_data$OverallAttitudeScore), ]
nrow(model_clean) ## Dropped ~1100 rows due to NA in AttitudeScore

# Select numeric variables
model_clean1 <- model_clean %>% 
  drop_na()
nrow(model_clean1)

```



```{r clean predictor factors, messages = FALSE, warning=FALSE}

# Check that empties have been dropped
summary(model_clean1$Education_short)

### Clean factor predictors for modelling
# Gender
model_clean2 <- model_clean1[model_clean1$Gender!= "N/A", ]
# Age
model_clean2 <- model_clean2[model_clean2$Age_short!= "N/A", ]
# Occupation
model_clean2 <- model_clean2[model_clean2$Occupation_short!= "Prefer not to say", ]
# Education
model_clean2 <- model_clean2[model_clean2$Education_short!= "Other", ]
# Drop empty factor levels
model_clean2 <- droplevels(model_clean2)
# Check n
nrow(model_clean2)
# Check that empties have been dropped
summary(model_clean2$Occupation_short)

### Manually categorised Occupation here to form new column = Occupation_short_clean
#### Also need to rename/ shorten some variable names
model_clean2 <- model_clean2 %>% 
  dplyr::rename(Attitude.Score = OverallAttitudeScore,
                Age = Age_short, 
                Proximity.to.release.site = SiteLocal,
                Urbanity.of.residence = Area_type,
                Survey.type = SurveyType,
                Occupation = Occupation_short_clean,
                Education = Education_short,
                Aware.of.white.storks = Q1_aware_stork,
                Heard.of.WSP = Q9_heard,
                General.reintroduction.support = Q14.5_agreement_score,
                Seen.white.stork = Q8_Seen,
                Time.in.nature = Q18_exp_nature,
                Visited.Knepp = Q27_Knepp_visit,
                Member.of.Environmental.Organisation = Q22_env_org_member, 
                Knowledge.score = KnowledgeScore,
                Bird.Interest.Score = BirdInterestScore,
                Env.Concern.Score = EnvConcern.score)

### Relevel Occupation factor 
model_clean2$Occupation <- relevel(model_clean2$Occupation,"Environment/Nature")
### Relevel Education factor 
model_clean2$Education <- relevel(model_clean2$Education,"No formal quals.")
### Relevel Seen(Wild/Captivity) factor 
model_clean2$Seen.white.stork <- factor(model_clean2$Seen.white.stork, levels = c("Wild", "Captivity", "No/Not sure"))
### Relevel Freq_exp_nature factor 
model_clean2$Time.in.nature <- relevel(model_clean2$Time.in.nature, "None") 

### Create two dataframes: one for each data collection for easier comparison
proact_data <- model_clean2[which(model_clean2$Survey.type == "Proactive"),]
natrep_data <- model_clean2[which(model_clean2$Survey.type == "NatRep"),]

```



### Exploring the response variable

**Response variable = Attitudes to WS reintroduction (Composite score)**

Now that the data has been cleaned we can explore the distribution of the response variable. We can see from the histogram (density plot) that the data is left-skewed as the reponses are geenrally towards the upper end of the response scale (0-5). A Shpiro test indicates that the distribution of non-normal, and QQ plots show that Squaring the response variable does the best job of normalising the distribution, but it's still non-normal.



```{r investigate response distribution, messages = FALSE, warning=FALSE}

str(model_clean2$Attitude.Score) # check it's a numeric column
model_clean2 %>%
     group_by(Survey.type) %>%
     summarise(sum(!is.na(Attitude.Score))) ## Counting sample size (Non-NA) values per survey type
  
## UNTRANSFORMED DATA
# QQ plot shows non-normality with skew towards the right
qqPlot(model_clean2$Attitude.Score)
# Also seen ina  density plot
ggdensity(model_clean2$Attitude.Score, 
          main = "Density plot of Attitude scores",
          xlab = "Overall Attitude Score")
mean(model_clean2$Attitude.Score)
median(model_clean2$Attitude.Score)
# Select variables
model_clean2 %>% 
  dplyr::select(Survey.type, Proximity.to.release.site, Attitude.Score) %>% 
  group_by()


# Running a Shapiro test to make sure
shapiro.test(model_clean2$Attitude.Score)

### TRANSFORMATIONS
# Square
qqPlot(model_clean2$Attitude.Score^2)
shapiro.test((model_clean2$Attitude.Score)^2)

# Exponential
qqPlot(exp(model_clean2$Attitude.Score))
# Inverse transformation for severse skew
qqPlot(1/(max(model_clean2$Attitude.Score+1) - model_clean2$Attitude.Score))

```




### Predictor correlation matrix of numeric variables

Moderate but insignificant correlation seen between all of the numeric score-based variables/predictors. BirdInterestScore tends to show strongest correlation with other predictors so this might be the most effective to remove from the model if the VIF score is also high.


```{r correlation matrix, messages = FALSE, warning=FALSE}

# Select numeric variables
model_numeric <- model_clean2 %>% 
  dplyr::select_if(., is.numeric) %>% 
  dplyr::select(., -UniqueID_all, -General.reintroduction.support) %>% 
  drop_na()
head(model_numeric)

# Create corrlation matrix
model.cor = cor(model_numeric, method = c("spearman"))
res1 <- cor.mtest(model_numeric, conf.level = .95)
# Create corrplot
corrplot::corrplot(model.cor, p.mat = res1$p, method = "circle", type = "upper", insig='blank',tl.col = "black",
         addCoef.col ='white', order = "AOE", diag=FALSE) ### All variables moderately correlated but not significant

```



### Global model generation

We now generate the global model. This is a saturated model with all of the fixed effects and their interesting interactions. There are no random effects in this model so we use a linear model.

```{r global glm model, messages = FALSE, warning=FALSE}

# Create saturated model (potentially SurveyType or locality as a random effect?)
global_model <- glm(Attitude.Score ~ Proximity.to.release.site + Survey.type + 
         Age + Gender + Urbanity.of.residence + Education + Occupation + Aware.of.white.storks + 
         General.reintroduction.support + Seen.white.stork +
         Visited.Knepp + # Have you visited Knepp?
         Time.in.nature + # Likert - experience of nature
         Heard.of.WSP + # Had you ever heard of the WSP and reintroduction of WS to southern England?
         Member.of.Environmental.Organisation + 
         Knowledge.score + NCI + ProCoBS + Bird.Interest.Score + Env.Concern.Score,
         data = model_clean2)

summary(global_model)        
par(mfrow = c(2, 2))
plot(global_model)
# http://www.sthda.com/english/articles/39-regression-model-diagnostics/161-linear-regression-assumptions-and-diagnostics-in-r-essentials/

# Check for variance inflation factors (VIF > 2 is worth removing and rechecking)
vif(global_model)
with(summary(global_model), 1 - deviance/null.deviance)

```

```{r new glm model, messages = FALSE, warning=FALSE}

### Create new model remoing Bird Interest Score and Heard of WSP
global_model1 <- glm(Attitude.Score ~ Proximity.to.release.site + Survey.type + 
         Age + Gender + Urbanity.of.residence + Education + Occupation + Aware.of.white.storks + 
         General.reintroduction.support + 
         Visited.Knepp + # Have you visited Knepp?
         Time.in.nature + # Likert - experience of nature
         Member.of.Environmental.Organisation + 
         Knowledge.score + NCI + ProCoBS + Bird.Interest.Score + Env.Concern.Score, data = model_clean2)
# Models summaries
summary(global_model1)
vif(global_model1)
# Check model residuals
par(mfrow = c(2, 2))
plot(global_model1) 

# Calculate r2
with(summary(global_model1), 1 - deviance/null.deviance)


## Sample sizes for the global model/dataset
c(table(model_clean2$SurveyType))
c(table(model_clean2$SiteLocal))

### Not in the model averaging = Freq.experience, SeenWildCaptivity, Heard of WSP, KnowledgeScore and BIS



```  



### Model selection

**Caveats to model selection**

* Depends on the models included in the candidate set.
* The parameter estimates and predictions arising from the “best” model or set of best models should be biologically meaningful.
* Need to decide whether to use model selection or common inferential statistics (e.g. based on P-values). Techniques that rely on both approaches are possible (e.g. backward variable selection followed by averaging of top models), such as the example provided above.


```{r top global model dredge, messages = FALSE, warning=FALSE}

#### MODEL SELECTION USING MUMIN PACKAGE
options(na.action = na.fail)

# Check VIF assumptions 
vif(global_model1)
# Dredge all possible models (model selection step)
attitude_dredge <- dredge(global_model1)
# Summarise the top model
summary(get.models(attitude_dredge, 1)[[1]])

#### Use a all-subsets model subsetting approach find a confidence set of models, recalculating weights each time:
attitude_del <- subset(attitude_dredge, delta <= 2, recalc.weights = TRUE) # delta(AIC) cutoff
# Save results table as a CSV file
# write.csv(as.data.frame(attitude_del), "Attitude_model_selection.csv")

```



### Model averaging

As we have so many predictors in the global model it’s unlikely that only one model explains all the variation in the data. From the dredge output we can see there is little difference in the AIC and weights of the first few models.

But how do we decide which model(s) to interpret? It's agreed that models with delta AIC (or other criterion) less than 2 are considered to be just as good as the top model, so therefore shouldn't be discounted. Additionally, we could use the weights: if a model has weight greater or equal to 95% then it is likely to be the top model. Otherwise we can generate a “credibility” set consisting of all models whose cumulative sum of AIC weights is 0.95. 

In any case, the point is that we have no good reason to exclude models other than the top one when the next models after it are likely to be just as good. Therefore, model averaging (AKA multi-model inference), is used to average the parameter estimates across multiple models and avoids the issue of model uncertainty. See below for the code and results of model averaging on this dataset for all models with a delta AIC <= 2.


**Key references**

* Harrison XA, Donaldson L, Correa-Cano ME, Evans J, Fisher DN, Goodwin CED, Robinson BS, Hodgson DJ, Inger R. 2018. A brief introduction to mixed effects modelling and multi-model inference in ecology. PeerJ 6:e4794 https://doi.org/10.7717/peerj.4794



```{r model averaging, messages = FALSE, warning=FALSE}

### Model averaging the top 10 models according to the delta AIC value
attitude_aves <- model.avg(get.models(attitude_del, subset = delta < 2))
summary(attitude_aves)


# View the model estimates
sjPlot::plot_model(attitude_aves, type = "est", vline.color = "black",   title = "",
                   order.terms = c(15,14,17,16,1,2,5,18,19,9,11,12,10,3,6,20,7,13,4,8), 
                   show.values = TRUE, value.offset = .3) + ylim(-.2, 0.5) + 
  scale_color_wsp("likert") + theme_sjplot()
ggplot2::ggsave(filename = "Attitude_averaging_table.png", width = 7, height = 8, dpi = 300)


# Print model averaged coefficients in table
attitude_confint <-as.table(round(confint(attitude_aves), 3))
attitude_exp <- as.table(exp(coefficients(attitude_aves)))
attitude_ave <- round(summary(attitude_aves)$coefmat.subset, 3)
# stargazer(attitude_ave, digits=3, title="Model averaged results", type = "html",
#           out="Attitude_averages_table.doc")

# Create tab_df table of model averaged estimates
attitude_ave1 <- tibble::rownames_to_column(as.data.frame(attitude_ave), "Predictor")
sjPlot::tab_df(attitude_ave1, title = "Model averaged results", alternate.rows = TRUE, digits=3, use.viewer = TRUE)


```




\newpage

### Exploring the relationship between attitude scores and Q15. "Do you support the reintroduction of WS to southern England?"


```{r lm attitudes Q15, messages = FALSE, warning=FALSE}

colnames(proact_data)
# How does attitude score vary by locality (within PROACTIVE dataset)
proact_data %>% 
  dplyr::select(Locality, Attitude.Score) %>% 
  drop_na() %>%  # N = 2,492
  dplyr::group_by(Locality) %>% 
  dplyr::summarise(
        n = n(),
        max_score = max(Attitude.Score, na.rm=T),
        mean_score = mean(Attitude.Score, na.rm=T),
        sd_score = sd(Attitude.Score, na.rm=T))
  
# Select variables and run model within a pipe %>% 
support_data <- proact_data %>% 
  dplyr::select(Locality, Attitude.Score, Support.for.WSP) %>% 
  drop_na()

# Running a ONE-way ANOVA
# QQ plot shows non-normality with skew towards the right
qqPlot(support_data$Attitude.Score)
model_support <- aov(Attitude.Score ~ Support.for.WSP, data = support_data)
summary(model_support)

# Create 
ggboxplot(support_data, x = "Support.for.WSP", y = "Attitude.Score",
          xlab = 'Support for white stork reintroductions', ylab = 'Attitude score')

```


